Equations & R Code

### 2016-06-24 manuscript/manuscript.Rmd
updates rel nonlin factorail plot to correct figure

\textbf{\large{Revisiting species coexistence reveals positive and negative relationships between species richness and ecosystem stability}}

@@

Model and Analysis
------------------
To discover how fluctuation-dependent coexistence mechanisms affect the diversity-stability relationship, we developed a semi-discrete, mechanistic consumer-resource model that allows coexistence of two species on one resource by either the storage effect or relative nonlinearity. 
In our model, the consumer can be in one of two-states: a dormant state $D$ and a live state $N$.
The dormant state could represent, for example, the seedbank of an annual plant.
Transitions between $N$ and $D$ occur at discrete intervals $\tau$ with continuous-time consumer-resource dynamics between discrete transitions. 
Thus, our model is formulated as "pulsed differential equations" [@Pachepsky2008; @Mailleret2009; @Mordecai2016]. 
For clarity we refer to $\tau$ as years and the growing time between years as seasons with daily ($t$) time steps.

@@

\noindent where $D(\tau_k)$, $N(\tau_k)$, and $R(\tau_k)$ are the abundances of each state at the end of growing season *k* and $\tau_k^+$ denotes the beginning of growing season $k=1$. 
The activation of dormant biomass to live biomass is controlled by $\gamma$, which is year (*k*) and species (*i*) specific.
Dormant biomass is equal to a constant fraction ($\alpha$) of live biomass at the end of the previous season ($N_{i}(\tau_k)$), plus survival ($1-\eta_i$) of dormant biomass ($D_{i}(\tau_k)$) at the end of the previous year and dormant biomass remaining after live biomass activation ($D_{i}(\tau_k)(1-\gamma_{i,\tau_k})$).
Live biomass is equal to newly activated dormant biomass ($\gamma_{i,t}[D_{i}(\tau_k)$), minus some fraction of live biomass that is converted to dormant biomass ($(1-\alpha_i)N_{i}(\tau_k)$)
We assume the resource pool is not replenished within a growing season.
Resource replenishment occurs between growing seasons, and the resource pool (*R*) at the start of the growing season *k*+1 is

\begin{align}
R(\tau_k^+) &= R^+,
\end{align}

\noindent $R^+$ is a random resource pulse drawn from a log-normal distribution with mean $\mu(R^+)$ and standard deviation $\sigma(R^+)$. Model parameters and notation are described in table 1.

@@

### Implementing the Storage Effect
To make this a ``storage effect" model, we need to satisfy three conditions: (1) the organisms must have a mechanism for persistence under unfavorable conditions, (2) species must respond differently to environmental conditions, and (3) the effects of competition on a species must be more strongly negative in good years relative to unfavorable years. 
Our model meets condition 1 because we include a dormant stage with very low death rates. 
We satisfy condition 2 with our model whenever $\gamma$ is not perfectly correlated between species. 
Lastly, our model meets condition 3 because condition 2 partitions intraspecific and interspecific competition into different years. 
Thus, during a high $\gamma$ year for one species, resource uptake is also inherently high for that species, which increases intraspecific competition relative to interspecific competition. 
So, given adequate variability in $\gamma$, the inferior competitor can persist. 
We created competitive hierarchies in the storage effect version of the model by altering species' biomass conversion efficiencies ($\epsilon$)

>>>>

\textbf{\large{Revisiting how species coexist reveals positive and negative relationships between species richness and ecosystem stability}}

@@

Model and Analysis
------------------
To discover how fluctuation-dependent coexistence mechanisms affect the diversity-stability relationship, we developed a semi-discrete consumer-resource model that allows many species to coexist on one resource by either the storage effect or relative nonlinearity. 
In our model, the consumer can be in one of two-states: a dormant state $D$ and a live state $N$.
The dormant state could represent, for example, the seedbank of an annual plant.
Transitions between $N$ and $D$ occur at discrete intervals $\tau$ with continuous-time consumer-resource dynamics between discrete transitions. 
Thus, our model is formulated as "pulsed differential equations" [@Pachepsky2008; @Mailleret2009; @Mordecai2016]. 
For clarity we refer to $\tau$ as years and the growing time between years as seasons with daily ($t$) time steps.

@@

\noindent where $D(\tau_k)$, $N(\tau_k)$, and $R(\tau_k)$ are the abundances of each state at the end of growing season *k* and $\tau_k^+$ denotes the beginning of growing season $k=1$. 
The activation of dormant biomass to live biomass is controlled by $\gamma$, which is year (*k*) and species (*i*) specific.
Dormant biomass is equal to a constant fraction ($\alpha$) of live biomass at the end of the previous season ($N_{i}(\tau_k)$), plus survival ($1-\eta_i$) of dormant biomass ($D_{i}(\tau_k)$) at the end of the previous year and dormant biomass remaining after live biomass activation ($D_{i}(\tau_k)(1-\gamma_{i,\tau_k})$).
Live biomass is equal to newly activated dormant biomass ($\gamma_{i,t}[D_{i}(\tau_k)$), minus some fraction of live biomass that is converted to dormant biomass ($(1-\alpha_i)N_{i}(\tau_k)$)
We assume the resource pool is not replenished within a growing season.
Resource replenishment occurs between growing seasons, and the resource pool (*R*) at the start of the growing season *k*+1 is $R(\tau_k^+) = R^+$, where $R^+$ is a random resource pulse drawn from a log-normal distribution with mean $\mu(R^+)$ and standard deviation $\sigma(R^+)$. 
Model parameters and notation are described in table 1.

@@

### Implementing the Storage Effect
To make this a storage effect model, we need to satisfy three conditions: (1) the organisms must have a mechanism for persistence under unfavorable conditions, (2) species must respond differently to environmental conditions, and (3) the effects of competition on a species must be more strongly negative in good years relative to unfavorable years. 
Our model meets condition 1 because we include a dormant stage with very low death rates. 
We satisfy condition 2 with our model whenever $\gamma$ is not perfectly correlated between species. 
Lastly, our model meets condition 3 because condition 2 partitions intraspecific and interspecific competition into different years. 
Thus, during a high $\gamma$ year for one species, resource uptake is also inherently high for that species, which increases intraspecific competition relative to interspecific competition. 
So, given adequate variability in $\gamma$, the inferior competitor can persist. 
We created competitive hierarchies in the storage effect version of the model by altering species' biomass conversion efficiencies ($\epsilon$)
